---
title: "r2weight"
author: "David Benkeser"
date: "August 22, 2016"
output: md_document
---
## Introduction

This package can be used to learn the convex combination of a multivariate outcome that maximizes cross-validated R-squared of a Super Learner-based prediction. Super Learner is an ensemble machine learning method with an associated [R package](https://github.com/ecpolley/SuperLearner) written by Eric Polley. The `SuperLearner` package provides a flexible interface for utilizing user-written wrappers to perform ensemble learning. A short tutorial on use of the Super Learner package can be found [here](http://benkeser.github.io/sllecture/).

In addition to estimating the best convex combination of outcomes for the sake of prediction, the `r2weight` package simultaneously estimates a function for predicting this learned convex combination and can compute a cross-validated estimate of the R-squared of this prediction algorithm for predicting the convex combination of outcomes. The package also computes influence function-based estimates of the standard error of the estimated performance metric that are used to construct confidence intervals and hypotheses tests. Finally, the package includes functions for computing variable importance measures for each variable. 

## Installation
The package can be installed directly from GitHub as follows:
```{r loadstuff, message=FALSE, warnings=FALSE}
# install/load devtools package
if(!("devtools" %in% row.names(installed.packages()))){
    install.packages("devtools")
}
library("devtools")

# install/load r2weight from GitHub
install_github("benkeser/r2weight")
library("r2weight")
```

## Use
The basic workflow of the package is to call the function `optWeight`, which estimates the optimal weights for the combined outcome. The S3-method `r2` can then be called on the `optWeight` object to obtain a cross-validated estimate of the R-squared for predicting the combined outcome. 

Here we illustrate the method using simulated data.
```{r callOptWeight, cache = TRUE, warnings = FALSE, message = FALSE}
# sample size
n <- 500

# set the seed
set.seed(12345)

# simulate nine covariate predictors
x1 <- runif(n,0,4)
x2 <- runif(n,0,4)
x3 <- runif(n,0,4)
x4 <- rbinom(n,1,0.75)
x5 <- rbinom(n,1,0.25)
x6 <- rbinom(n,1,0.5)
x7 <- runif(n,0,4)
x8 <- runif(n,0,4)
x9 <- runif(n,0,4)
# put all predictors in single data.frame
X <- data.frame(x1=x1, x2=x2, x3=x3, x4=x4, x5=x5, 
                x6=x6, x7=x7, x8=x8, x9=x9)

# simulate three outcomes
y1 <- x1 + 2*x2 + 4*x3 + x4 + 2*x5 + 4*x6 + 2*x7 + rnorm(n, 0, 5)
y2 <- x1 + 2*x2 + 4*x3 + x4 + 2*x5 + 4*x6 + 2*x8 + rnorm(n, 0, 5)
y3 <- x1 + 2*x2 + 4*x3 + x4 + 2*x5 + 4*x6 + 2*x9 + rnorm(n, 0, 5)
# put all outcomes in single data.frame
Y <- data.frame(y1 = y1, y2 = y2, y3 = y3)

# call optWeight using simple Super Learner library
out1 <- optWeight(Y = Y, X = X, SL.library = c("SL.mean","SL.glm","SL.step"))

# print the object
out1
```

The estimated optimal weights are shown (note that the true weights in this example are 1/3, 1/3, 1/3) along with the estimated cross validated R-squared for predicting each outcome using the Super Learner. 
